\LARGE \bf Face Recognition: Too Bias, or Not Too Bias?

### 2020-02-16 sections/analysis.tex
Version for arXiv

\section{Results and Analysis}
An {\em off-the-shelf} \gls{cnn} was used throughout to control the model across all experiments. For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped. The two features were fused by average pooling (\ie 512 D).

@@

\begin{figure}[t!] 
	\centering    
	\includegraphics[width=\linewidth] {figures/human_eval.pdf}
		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}. Accuracy scores shown as bar plots. Humans are most successful at recognizing their own subgroup, with a few exceptions (\eg bottom).}}
		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
		\label{fig:human-eval} 
% 		\vspace{-1mm}
\end{figure} 
% \vspace{-4mm}
% \vspace{-5pt}
\subsection{Human evaluation}
 Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectfully. Subjects of a subgroup likely have mostly been exposured to others of the same. Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but other gender. Our findings concur. Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex. Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-to-best having the same ethnicity and opposite gender. On average, subgroups are comparable at labeling images.

>>>>

\section{Results and Analysis}
A single \gls{cnn} was used as a means to control the experiments. For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about a 13\%, 14\%, 3\%, and 70\% ratio in Asians, Blacks, Indians, and Whites, respectively; furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2019racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.

@@

\begin{figure}[t!] 
	\centering    
	\includegraphics[width=\linewidth] {figures/human_eval.pdf}
		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}. Accuracy scores shown as bar plots. Humans are most successful at recognizing their own subgroup, with a few exceptions (\eg bottom).}}
		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
		\label{fig:human-eval} 
% 		\vspace{-1mm}
\end{figure} 
% \vspace{-4mm}
% \vspace{-5pt}
\subsection{Human evaluation}
 Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectfully. Subjects of a subgroup likely have mostly been exposed to others of the same. Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but other gender. Our findings concur. Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex. Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-to-best having the same ethnicity and opposite gender. On average, subgroups are comparable at labeling images. Analogous to the \gls{fr} system, performance ratings differ when analyzing within and across subgroups. In other words, performance on \gls{bfw} improved with subgroup-specific thresholds. Similarly, humans tend to better recognize individuals by facial cues of the same or similar demographics. Put differently, as the recognition performances drops with a global threshold optimized for one subgroup and deployed on another, human performance tends to drop when across subgroups (\ie most familiar with the same subgroup; performance seems to drop when faced with another, less familiar subgroup).

%  for it better adapts wh specific to subgroup) shift in sensitivity across subgroups when deep encodings are compared, the underlying mechanism

### 2020-02-16 sections/section3.tex
Version for arXiv

\vspace{1mm}
\noindent\textbf{Validating labels.} 
Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}. A matrix of cosine similarity scores was then generated for each subject, and removed samples (\ie rows) with median scores below threshold $\theta=0.2$. Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, with the percentile $P=50$, fell below a threshold of $\theta=0.2$ (set manually). This allowed us to quickly prune \gls{fp} face detections. Following~\cite{robinson2016families, robinson2018visual}, we built a JAVA tool to visually validate the remaining faces. For this, the faces were ordered by decreasing confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window. Labeling then consisted of going subject-by-subject and flagging faces of \emph{imposters}.

>>>>

\vspace{1mm}
\noindent\textbf{Validating labels.} 
Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}: applying an affine transformation to align faces according to pre-defined eye locations, each face fed through twice (\ie the original and horizontally flipped), with the final encoding obtained by fusing the two features by average pooling (\ie 512 D). A matrix of cosine similarity scores was then generated for each subject, and removed samples (\ie rows) with median scores below threshold $\theta=0.2$ (set manually). Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, where $P=50$. In other words, the median (\ie 50 percentile) of all scores for a faces with respect to all of faces for the respective subject must pass a threshold of $\theta=0.2$; else, the face is dropped. This allowed us to quickly prune \gls{fp} face detections. Following~\cite{robinson2016families, robinson2018visual}, we built a JAVA tool to visually validate the remaining faces. For this, the faces were ordered by decreasing confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window. Labeling then consisted of going subject-by-subject and flagging faces of \emph{imposters}.
